\vspace{-3em} \textbf{ Mapping stellar surfaces\\ II: An interpretable Gaussian process model for light curves }

### 2021-01-28 paper2/ms.tex
working in intro to paper2

A major source of stellar variability in both light curves and radial velocity
datasets is the modulation induced by surface features like starspots
rotating in and out of view. While GPs excel at marginalizing over stellar rotational
variability, they have been less useful when the goal is to make
inferences about the actual source of this variability, such as the
properties of starspots and the magnetic processes that generate them.
\textbf{While it is straightforward to derive
    posterior constraints on the hyperparameters of an effective GP model for
    observations of a star, it is not clear what those constraints actually
    tell us about the stellar surface.}
Specifically, in all but a few restricted cases, there is no \emph{first principles}
relationship between the descriptive parameters of a typical GP model
(see \S\ref{sec:gp-intro})
and the physical properties of the stellar surface that is being observed.
%
For instance, it may be tempting to interpret the GP amplitude hyperparameter
as some measure of the spot contrast or the total number of spots,
or the GP timescale hyperparameter as the
spot lifetime, but there are no guarantees these interpretations will
hold in general. After all, the choice of kernel is quite often \emph{ad hoc},
providing an \emph{effective}---as opposed to \emph{interpretable}---description
of the physics.
%
There are two important exceptions to this: asteroseismic studies, in which the
the GP hyperparameters can offer direct insight into
the behavior of complex pulsation modes and thus physical properties of the
stellar interior \citep[e.g.,][]{BrewerStello2009,ForemanMackey2017}; and
stellar rotation period studies, in which the period hyperparameter $P$ can usually
be associated with the rotation period of the star \citep[e.g.,][]{Angus2018}.%
\footnote{An exception to this is in the presence of strong differential
    rotation, in which case many periods may be present in the data, or
    when spots evolve coherently, which can also introduce weak periodicities
    in the light curve.}
%
For spot-induced variability, on the other hand,
GPs are usually used
when the variability itself is a nuisance parameter. For example, if the
goal is to constrain the properties of a transiting exoplanet or to
search for a planetary signal in a radial velocity dataset, a GP might be
used to remove (or, better yet, to marginalize over) the stellar variability
\citep[e.g.,][]{Haywood2014,Rajpaul2015,Luger2017b}.
In this case, the physics behind the variability is irrelevant, so an
effective model of this sort is generally sufficient.

Understanding the properties of starspots, faculae, plages, and other
features on the surfaces of stars is crucial to \xxx{UNDERSTANDING MAGNETIC
    FIELDS, STELLAR INTERIORS, MITIGATING EXOPLANET SYSTEMATICS} \citep{Rackham2018}.

@@

We would like to thank David W. Hogg, Michael Gully-Santiago, and the
Astronomical Data Group at the Center for Computational Astrophysics for
many thought-provoking discussions that helped shape this paper.

>>>>

A major source of stellar variability in both light curves and radial velocity
datasets is the modulation induced by magnetically-driven surface features like starspots
rotating in and out of view. While GPs excel at marginalizing over stellar rotational
variability, they have been less useful when the goal is to make
inferences about the actual source of this variability, such as the
properties of starspots and the magnetic processes that generate them.
\textbf{While it is straightforward to derive
    posterior constraints on the hyperparameters of an effective GP model for
    observations of a star, it is not clear what those constraints actually
    tell us about the stellar surface.}
Specifically, in all but a few restricted cases, there is no \emph{first principles}
relationship between the descriptive parameters of a typical GP model
(see \S\ref{sec:gp-intro})
and the physical properties of the stellar surface that is being observed.
%
For instance, it may be tempting to interpret the GP amplitude hyperparameter
as some measure of the spot contrast or the total number of spots,
or the GP timescale hyperparameter as the
spot lifetime, but there are no guarantees these interpretations will
hold in general. After all, the choice of kernel is quite often \emph{ad hoc},
providing an \emph{effective}---as opposed to \emph{interpretable}---description
of the physics.
%
There are two important exceptions to this: asteroseismic studies, in which the
the GP hyperparameters can offer direct insight into
the behavior of complex pulsation modes and thus physical properties of the
stellar interior \citep[e.g.,][]{BrewerStello2009,ForemanMackey2017}; and
stellar rotation period studies, in which the period hyperparameter $P$ can usually
be associated with the rotation period of the star \citep[e.g.,][]{Angus2018}.%
\footnote{An exception to this is in the presence of strong differential
    rotation, in which case many periods may be present in the data, or
    when spots evolve coherently, which can also introduce weak periodicities
    in the light curve.}
%
For spot-induced variability, on the other hand,
GPs are usually used
when the variability itself is a nuisance parameter. For example, if the
goal is to constrain the properties of a transiting exoplanet or to
search for a planetary signal in a radial velocity dataset, a GP might be
used to remove (or, better yet, to marginalize over) the stellar variability
\citep[e.g.,][]{Haywood2014,Rajpaul2015,Luger2017b}.
In this case, the physics behind the variability is irrelevant, so an
effective model of this sort is generally sufficient.

However, understanding the properties of stellar surfaces and starspots
in particular is a crucial step
toward understanding stellar magnetism, which plays a fundamental part
in stellar interior structure and evolution. Stellar magnetic fields
control the spin-down of stars over time, on which the field of
gyrochronology is founded \citep[][]{Barnes2001,Angus2019}. They affect
wave propagation in stellar interiors and must be properly
understood to interpret asteroseismic measurements \citep[e.g.,][]{Fuller2015}.
Strong magnetic fields are also likely the driving force behind chemical
peculiarity in Ap/Bp stars~\citep{Turcotte2003,Sikora2018}, as well as
radius inflation in M dwarfs \citep{Gough1966,Ireland2018}.
Significant effort has thus been spent on theoretical and numerical approaches to understanding
magnetohydrodynamics (MHD) in stars~\citep[e.g.,][]{Miesch2009}.
Theoretical progress is challenging; MHD equations are nonlinear, since
magnetism often emerges in regions of highly developed turbulence. There is a
large separation of scales between the size of a star and the
magnetic/thermal/momentum diffusion length-scales, and so numerical
simulations at realistic resolutions remain out of reach. As a result
there is still significant theoretical uncertainty over how dynamos
manifest in stars of different masses, how magnetism interacts with rotation,
and how star spots form \citep{Yadav2015,Weber2016}. Because of the
significant impacts of magnetism on every branch of stellar astrophysics,
these difficulties make new constraints from observations especially valuable.

\xxx{MITIGATING EXOPLANET SYSTEMATICS} \citep{Rackham2018}.

@@

We would like to thank David W. Hogg, Michael Gully-Santiago, Adam Jermyn, and the
Astronomical Data Group at the Center for Computational Astrophysics for
many thought-provoking discussions that helped shape this paper.
